---
title: "Data Visualization"
author: "Paloma Cartwright"
date: "18/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)


library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library(sf)

```
# Working with lizards
## Get the lizards data 

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

## Make a really bad histogram of lizard weights 
```{r}
ggplot(lizards, aes(x = weight)) +
  geom_histogram(fill = "orange", 
                 color = "black", 
                 size = 0.2, #width of the border 
                 linetype = "dotted")
```
## Create a less ugly graph 

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(shape = 22, 
             size = 3,
             alpha = 0.5,
             aes(color = common_name))

```

## Make a very ridiculous, bad idea graph 

Bubble plot where the color of the points changes based on the common_name and the size of points varies based on the total_length 

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, 
                 size = total_length), 
             shape = 22, 
             fill = "black") + 
  theme_minimal()

```


## Create something not quite as terrible 

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name), 
             size = 0.8) +
  theme_light() + 
  facet_wrap(~common_name, ncol = 4)
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name), 
             size = 0.8) + 
  facet_grid(sex ~ tail)
```

## I don't even know what at this point 

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# Same thing: but it is better to use summarize if you also need to do mean, std and other stuff 
#lizard_counts <- lizards %>% dplyr::count(common_name)

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
```


```{r}

ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, -count), x = count)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(title = "Total Number of Lizard by Species", 
       y = "Common Name", 
       x = "Total Number of Lizards")
  

```

Try converting common name to an ordered factor outside of ggplot

```{r}
lizard_counts <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))

```


## Axis scale breaks, limits & labels 


Scatter plot 
```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  scale_x_continuous(breaks = c(0, 5, 50, 500), 
                     limits = c(0, 500), 
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())


```

```{r}
ggplot(data = lizard_counts, aes(x = date, y = count)) +
  geom_line(aes(color = count), show.legend = FALSE) +
  scale_x_date(date_breaks = "3 years", 
               date_labels = "%y")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple"), 
                     breaks = c(0, 20, 40))
```

### Update a color scheme using a palette in paletteer

Make a horizontal boxplot with common_name on the y_axis, total_length on the x, with color changing based on common_name 

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))


ggplot(data = lizards_fct, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "khroma::bright")

```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid = element_blank(), 
    plot.background = element_rect(fill = "cyan4"), 
    axis.text.x = element_text(color = "orange"), 
    axis.title = element_text(color = "yellow", size = 10)
  )
```
# Day 9 Work: Part 2 of data visualization

## ggrepel

```{r}

ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", site == "sand")

ggplot(data = ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num), size = 2, max.overlaps = 20)

```

## Using gapminder dataset 

```{r}
gapminder_Europe <- gapminder %>% 
  filter(continent == "Europe",  year == "2007")

ggplot(data = gapminder_Europe, aes(x= gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)
```

## gghighlight

```{r}

p <- ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)
```

```{r}

q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) +
  gghighlight(max(weight) > 30)

q 

```

## Patchwork for compoint figures 

```{r}
p / q & # p is on top of q 
  theme_minimal()

p | q #these are side by side 

```
## A few new graph types 

### Marginal plots 

```{r}

whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_rug()

```

Marginal plot with boxplots 

```{r}

my_plot <- ggplot(whiptails, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2, shape = 22) +
  scale_color_manual(
    values = c("cyan3", "black", "gold"),
    name = "Sex:",
    labels = c("Female", "Juvenile", "Male")
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggMarginal(my_plot, type = "boxplot", groupColour = TRUE)


```

### Beeswarm plot

```{r}

ggplot(data = whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex)) +
  geom_boxplot(fill = NA)

```

### Heatmap with geom_tile()

```{r}

lizards_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(year = lubridate::year(date), common_name) %>% 
  summarize(count = n())

```

```{r}

ggplot(data = lizards_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = count)) +
  geom_text(aes(label = count), color = "white", size = 2) +
  scale_fill_viridis_c()
  
```


### Make a map 

```{r}

jornada_vegetation <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()

ggplot(data = jornada_vegetation) +
  geom_sf(aes(fill = name), color = "white", size = 0.2) +
  scale_fill_paletteer_d("ggthemes::manyeys") +
  labs(x = "Longitude", 
       y = "Latitude", 
       fill = "Dominant Vegetation") +
  theme_minimal()

```

