---
title: "Data Visualization"
author: "Paloma Cartwright"
date: "18/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)


library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library(sf)

```

## Get the lizards data 

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

## Make a really bad histogram of lizard weights 
```{r}
ggplot(lizards, aes(x = weight)) +
  geom_histogram(fill = "orange", 
                 color = "black", 
                 size = 0.2, #width of the border 
                 linetype = "dotted")
```
## Create a less ugly graph 

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(shape = 22, 
             size = 3,
             alpha = 0.5,
             aes(color = common_name))

```

## Make a very ridiculous, bad idea graph 

Bubble plot where the color of the points changes based on the common_name and the size of points varies based on the total_length 

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, 
                 size = total_length), 
             shape = 22, 
             fill = "black") + 
  theme_minimal()

```


## Create something not quite as terrible 

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name), 
             size = 0.8) +
  theme_light() + 
  facet_wrap(~common_name, ncol = 4)
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name), 
             size = 0.8) + 
  facet_grid(sex ~ tail)
```

## I don't even know what at this point 

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# Same thing: but it is better to use summarize if you also need to do mean, std and other stuff 
#lizard_counts <- lizards %>% dplyr::count(common_name)

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
```


```{r}

ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, -count), x = count)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(title = "Total Number of Lizard by Species", 
       y = "Common Name", 
       x = "Total Number of Lizards")
  

```

Try converting common name to an ordered factor outside of ggplot

```{r}
lizard_counts <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))

```


## Axis scale breaks, limits & labels 


Scatter plot 
```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  scale_x_continuous(breaks = c(0, 5, 50, 500), 
                     limits = c(0, 500), 
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())


```

```{r}
ggplot(data = lizard_counts, aes(x = date, y = count)) +
  geom_line(aes(color = count), show.legend = FALSE) +
  scale_x_date(date_breaks = "3 years", 
               date_labels = "%y")
```

